---
title: "Stats 744: hw2"
author: "Daniella Lato"
date: "20/09/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggthemes)
library(ggplot2)
library(directlabels)
options(scipen=10000)
```

## Original Graph
The following homework assignment is about Jia You’s online graphic about vaccines which can be found
[here.](https://www.sciencemag.org/news/2017/04/here-s-visual-proof-why-vaccines-do-more-good-harm)


## What question is this graphic trying to answer?
I believe that this graphic is trying to answer the question: Do vaccines work?
I think that the author is trying to show that after a vaccine has been licensed for a number of diseases, the number of reported cases (in the U.S.) decreases.
Hence, matching the title of the graphic that "visual proof of why vaccines do more good than harm".

## Create 2 new plots with the same data

Read in and prep the data.
```{r read dat}
vac_dat <- read.csv("vaccine_data_online.csv")
#now re-ordering the data 
vac_dat$disease <- with(vac_dat, reorder(disease, cases, FUN = mean))
vac_dat$disease <- with(vac_dat, reorder(disease, cases, FUN =function(x) -x[1]))
#make dataframe for vaccination licenceing
vac_l <- vac_dat[which(vac_dat$vaccine != "FALSE"),]
vac_l <- vac_l[,c(3,4)]
vac_l <- vac_l[!duplicated(vac_l$disease),]
```

### Graph 1:

The first graph shows the total number of reported cases for each disease in the U.S.
The black vertical lines show the year that the vaccine was licensed for each disease.
```{r g1}
g1 <- ggplot(vac_dat, aes(x = year, y = cases, color = disease))+
  geom_line()+
  theme_classic() +
  theme(legend.position = "none")
g2_fac <- g1 + facet_wrap(~disease)
g2_fac + geom_vline(data = vac_l, aes(xintercept = year)) +
  xlab("Year") +
  ylab("Number of Reported Cases in U.S.")
```


Log Transformed:

The graph below is the same as the previous one but with the number of cases has been log transformed (log base 10).
I found it hard to notice the decrease in the number of reported cases in some of the diseases like Pertussis and Polio because there were less reported cases than Measles.

```{r g1log}
#log transform
vac_dat$cases <- log10(vac_dat$cases)
#graph
g1_log <- ggplot(vac_dat, aes(x = year, y = cases, color = disease))+
  geom_line()+
  theme_classic() +
  theme(legend.position = "none")
g2_log_fac <- g1_log + facet_wrap(~disease)
g2_log_fac + geom_vline(data = vac_l, aes(xintercept = year)) +
  xlab("Year") +
  ylab("Number of Reported Cases in U.S. (log10)")
```

### Discussion:

In all of the above graphs the data is read from right to left which may be culturally easier to read than the original graph which is read from top to bottom.
I also find it easier to see the decreasing trend when it is presented as a line.
Classically humans are better at deciphering trends when they are on a common axis, which is what my graphs show compared to the original graph by Jia You.
Area is also ranked low on Cleavland's list of how accurately humans can visually process data.
I think this makes it hard to interpret Jia You's graph because you are forced to decipher area instead of position along a common axis.



### Graph 2:
This graph shows the same vaccination data as distributions on a ridgeline graph. The black vertical lines represent the year that the vaccine was licenced for the disease.
The number of cases has been log transformed.

```{r g2}
g_area <- ggplot(vac_dat, aes(x = year, y = cases)) + 
  geom_area(aes(color = disease, fill = disease), 
            alpha = 0.5, position = position_dodge(0.8))+
  theme_classic() +
  theme(legend.position = "none") +
  geom_vline(data = vac_l, aes(xintercept = year)) +
  xlab("Year") +
  ylab("Number of Reported Cases in U.S. (log10)")
  
g_area + facet_wrap(~disease)
```


### Discussion:
This graph utalizes the area component of Jia You's graph, having a physical area of the graph represent the number of cases.
I find it easier to see the trend that the number of cases (in the U.S.) decreases after the vaccination date with the area "coloured in".
It is easy to see that at the beginning of the time series (before vaccination licence) that the number of cases was higher than after the vaccination of each disease was licenced.
It is also helpful to have all the graphs displayed in one picture in an orderly format.
This makes it easy to compare between diseases and identify similar trends across all diseases.



## Fun Graphic:
This homework reminded me of one of my favorite data visualization artists Mona Chalabi. She has a really neat visualization of vaccination data
[here.](https://instagram.com/p/BtTjIYwngqn/)