---
title: "Stats744: HW4"
output: pdf_document
fontsize: 12pt
header-includes: \usepackage{xspace, longtable, booktabs}
---

```{r setup, include=FALSE}
library(dotwhisker)
library(broom)
library(dplyr)
library(tidyverse)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(ggplot2)
library(flextable)
options(scipen=10000)
theme_set(theme_bw() + theme(strip.background =element_rect(fill="#e7e5e2")) +
            theme(strip.text = element_text(size =10)) +
            theme(plot.title = element_text(hjust = 0.5), 
                  panel.background = element_rect(fill = "white", colour = NA), 
                  panel.grid.major = element_line(colour = "grey90", size = 0.2), 
                  panel.grid.minor = element_line(colour = "grey98", size = 0.5), 
                  panel.spacing = unit(0.25, "lines"), 
                  axis.text=element_text(size=10),
                  legend.position="top") 
          )
```


\newcommand{\smel}{\textit{S.\,meliloti}\xspace}
\newcommand{\bass}{\textit{B.\,subtilis}\xspace}
\newcommand{\ecol}{\textit{E.\,coli}\xspace}
\newcommand{\strep}{\textit{Streptomyces}\xspace}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}



## Illustrating Statistical Inference

### Data:

As part of my thesis I have information about substitutions and their variation across four bacterial genomes: \ecol, \bass, \strep, and \smel.
The bacteria \smel
is a multi-repliconic bacteria and therefore each of it's replicons is analized separatly. 
The data is binary, stating at each site in the genome if there is a substitution, or there is not.
This was also broken down into protein coding and non-protein coding sections of the genome.
I previously fit a logistic regression to the data and obtained a table with the following results:


\begin{table}[h]
	\centering
		\begin{tabular}{lcccc}
			\toprule
			Bacteria and Replicon & Coefficent Estimate & Standard Error & z-value & P($>$|z|)\\
			\midrule
			\ecol Chromosome &  -4.308\e{-8} & 2.584\e{-9} & -16.67 & $<$2\e{-16} \\
			\bass Chromosome &  -4.971\e{-8} & 4.268\e{-9} & -11.65 & $<$2\e{-16} \\
			\strep Chromosome & 1.989\e{-8} & 8.696\e{-10} & 57.37 & $<$2\e{-16}\\
			\smel Chromosome &  -1.903\e{-7} &  2.13\e{-8} & -8.934 & $<$2\e{-16} \\
			\smel pSymA &  -6.642\e{-7} & 2.801\e{-8} & -23.71 & $<$2\e{-16} \\
			\smel pSymB & 1.769\e{-7} & 2.33\e{-8} & 7.593 & 3.11\e{-14}\\
			\bottomrule
		\end{tabular}
\end{table}


\begin{table}[h]
	\centering
		\begin{tabular}{lcccc}
			\toprule
			Bacteria and Replicon & Coefficent Estimate & Standard Error & z-value & P\\
			\midrule
			 \ecol Chromosome &  9.896\e{-9} & 7.159\e{-9} & 1.382 & 0.167 \\
			 \bass Chromosome &  -1.055\e{-7} & 1.25\e{-8} & -8.436 & $<$2\e{-16}  \\
			 \strep Chromosome & 1.635\e{-7} & 2.893\e{-9} & 56.5 & $<$2\e{-16}\\
			 \smel Chromosome & -2.900\e{-7} & 3.852\e{-8} &  -7.527 & 5.18\e{-14} \\
			 \smel pSymA &  -1.263\e{-6} & 5.595\e{-8} & -22.57 & $<$2\e{-16} \\
			 \smel pSymB & 4.771\e{-7} & 5.314\e{-8} & 8.978 & $<$2\e{-16}  \\
			\bottomrule
		\end{tabular}
\end{table}


Read in the data and format: (I have over 160 million lines of data so it was just not feisable to run the logistic regressions. So I manually added the output from the logistic regression results)

I think I may have cheated a bit, but I made each bacteria a "term" even though they were not actually "terms" in the model.
Each logistic regression model was run separatly on both the protein coding and non-protein coding sections of each bacterial replicon.
```{r}
subs_dat <- read.table(header=TRUE,check.names=FALSE,text="
term model  estimate  std.error statistic p.value 
ecoli 1 -0.00000004308 0.000000002584  -16.67 0.0000000000000002
ecoli 0 0.000000009896  0.000000007159  1.382 0.167
bass  1 -0.00000004971  0.000000004268  -11.65  0.0000000000000002
bass  0 -0.0000001055 0.0000000125  -8.436  0.0000000000000002
strep  1 0.00000001989 0.0000000008696 57.37 0.0000000000000002
strep 0 0.0000001635  0.000000002893  56.5  0.0000000000000002
schrom  1 -0.0000001903 0.0000000213  -8.934  0.0000000000000002
schrom  0 -0.0000002900 0.00000003852 -7.527  0.0000000000000518
pa  1  -0.0000006642  0.00000002801 -23.71  0.0000000000000002
pa  0 -0.000001263  0.00000005595 -22.57  0.0000000000000002
pb  1 0.0000001769  0.0000000233  7.593 0.0000000000000311
pb  0 0.0000004771  0.00000005314 8.978 0.0000000000000002

")
```

### Graph
```{r}
dwplot(size = 3, subs_dat) %>%
  relabel_predictors(ecoli = "E.coli",
                          bass = "B. subtilis",
                          strep = "Streptomyces",
                          schrom = "S.meliloti Chromosome",
                          pa = "S.meliloti pSymA",
                          pb = "S.meliloti pSymB", face = "italic") +
  xlab("Coefficent Estimates") +
  geom_vline(xintercept=0,lty=2)+ 
  scale_colour_manual(values=c("#E08F8F","#173FAA"),
                      name = "",
                      labels = c("Protein Coding", "Non-Protein Coding"))
 
```


### Discussion



I was really excited for this assignment because I think it is a cool way to present multiple coefficent estimates focusing on their significance.
It never occured to me to present my logistic regression results like this.

```{r}
#I tried many things to make the bacteria names italic (like the below code), but I just could not get it to work.
#dwplot() appears to not like expressions or paste...or substitute
#subs_dat %>% dwplot(relabel_predictors(c("ecoli" = expression(paste(italic("E.coli"), " Chromosome")),
#                          "bass" = expression(paste(italic("B. subtilis"), " Chromosome")),
#                          "strep" = expression(paste(italic("Streptomyces"), " Chromosome")),
#                          "sinoC" = expression(paste(italic("S.meliloti"), " Chromosome")),
#                          "pSymA" = expression(paste(italic("S.meliloti"), " pSymA")),
#                          "pSymB" = expression(paste(italic("S.meliloti"), " pSymB"))))) +
#  xlab("CHANGE ME (x10^2)") +
#  geom_vline(xintercept=0,lty=2)+ 
#  scale_colour_manual(values=c("#E08F8F","#173FAA"),
#                      name = "",
#                      breaks=c("0","1"),
#                      labels = c("Protein Coding", "Non-Protein Coding"))

#I also tried this before I made the dwplot
##make sure bacteria names are italic
#bac_names <- c("ecoli" = expression(paste(italic("E.coli"), " Chromosome")),
#                          "bass" = expression(paste(italic("B. subtilis"), " Chromosome")),
#                          "strep" = expression(paste(italic("Streptomyces"), " Chromosome")),
#                          "sinoC" = expression(paste(italic("S.meliloti"), " Chromosome")),
#                          "pSymA" = expression(paste(italic("S.meliloti"), " pSymA")),
#                          "pSymB" = expression(paste(italic("S.meliloti"), " pSymB")))
# 
#subs_dat$term <- bac_names
 
```


